### 2019-09-20 MoneyBallR_RMD.Rmd
EDA and preliminary feature engineering

---
title: "MoneyBallR_RMD"
author: "Corey Arnouts"
date: "9/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Decision Trees to Find Variable Interactions
```{r}
library(tidyverse)

df_train <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-training-data.csv")

df_test <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-evaluation-data.csv")

source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")

colnames(df_train) <- sub("TEAM_", "", colnames(df_train))

colnames(df_test) <- sub("TEAM_", "", colnames(df_test))

df_test <- RF_with_Nulls(df_train,df_test,"TARGET_WINS",.5,10,25,.02,5,1)

```

# See if we can impute some misssing values Using a Random Forest
```{r}
library(randomForest)
df_train.na <- df_train
df_train.imputed <-rfImpute(TARGET_WINS ~.,df_train.na)

str(df_train)
str(df_train.imputed)

df_train <- df_train.imputed
```

#Correlation Matrix
```{r}

na_count <-sapply(df_train, function(y) sum(length(which(is.na(y)))))

na_count <- data.frame(na_count)

library(corrplot)

M <- cor(df_train)
corrplot(M, method = "circle") #plot matrix

library("PerformanceAnalytics")

chart.Correlation(df_train, histogram=TRUE, pch=19)
```

#Principal Components
```{r}
#PCA Analysis in R DataCamp Article
library(tidyverse)

mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE) #exclude the categorical variables

mtcars <- mtcars

summary(mtcars.pca)

str(mtcars.pca)

library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)

ggbiplot(mtcars.pca) # this shows you what attributes contribute to principal component 1 and 2

ggbiplot(mtcars.pca, labels=rownames(mtcars)) # add in the sample information

mtcars.country <- c(rep("Japan", 3), rep("US",4), rep("Europe", 7),rep("US",3), "Europe", rep("Japan", 3), rep("US",4), rep("Europe", 3), "US", rep("Europe", 3))

ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country) #represent circles around groupings I want to do this binary classifiers

ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components

################################################################################
##################### Same thing but with MoneyBallR data ######################
################################################################################

df_train.pca <- prcomp(df_train[,c(3:17)], center = TRUE,scale. = TRUE) #exclude the categorical variables

df_train.target <-

summary(df_train.pca)

str(df_train.pca)

library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)

ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)

ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country) # take a look at some of the other principal components

library(tidyverse)

test <- df_train %>% dplyr::select(BATTING_H)

df_train.pca <- df_train %>% dplyr::select(PITCHING_HR,BATTING_HR) %>%  prcomp(center = TRUE,scale. = TRUE) #exclude the categorical variables

summary(df_train.pca)

str(df_train.pca)

library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)

ggbiplot(df_train.pca) + geom_point(size = .01)+ xlim(-8,8)+ ylim(-10,5)

ggplot(df_train, aes(x=PITCHING_HR, y=BATTING_HR, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles

df_train$HR_RATIO <- df_train$BATTING_HR/df_train$PITCHING_HR

library(plotly)
ggplotly(ggplot(df_train, aes(x=PITCHING_HR, y=HR_RATIO, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
scale_colour_gradient(low="cyan", high="firebrick1")     # Use hollow circles
)

mid <- mean(df_train$TARGET_WINS)

ggplotly(ggplot(df_train, aes(x=BATTING_H, y=PITCHING_H, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" )     # Use hollow circles
)

## look at batting hits to pitching hits ratio
df_train$Hitting_Ratio <- df_train$BATTING_H/df_train$PITCHING_H

ggplotly(ggplot(df_train, aes(x=BATTING_H, y=Hitting_Ratio, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" )     # Use hollow circles
)

for (i in 1:ncol(df_train)){
ggplotly(ggplot(df_train, aes(x=df_train[[i]], y=TARGET_WINS, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(low = "blue",
                            high = "red", space = "Lab" )     # Use hollow circles
)

}

ggplotly(ggplot(df_train, aes(x=BATTING_2B , y=BATTING_H, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(midpoint = mid, low = "blue", mid = "white",
                            high = "red", space = "Lab" ) )

```

#Predict Hits Batted by Hits Pitched and then see what the difference is so like hits batted above expectation
And then use this as a predictive feature

```{r}
#First Fit Target Wins just by Batted_H and Pitched_H

#model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = df_train)

#df_train$Prediction <- predict(model,df_train)

#summary(model)
## .45 correlation 
##try on it on out of sample

### Add my meta features ###
model <- lm(BATTING_H ~ PITCHING_H, data = df_train)

df_train$BATTING_H_EXP <- predict(model,df_train)
df_train$BATTING_H_ABV_EXP <- df_train$BATTING_H/df_train$BATTING_H_EXP

train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]

model <- lm(TARGET_WINS ~ BATTING_H + PITCHING_H, data = train_data)

hold_out_data$Prediction <- predict(model,hold_out_data)

cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)

summary(model)

############# FITTING THE MODEL WITH EVERYTHING ###################
############# BASELINE ############################################
df_train$INDEX <- NULL

train.index1 <- createDataPartition(df_train$TARGET_WINS, p = .7, list = FALSE)
train_data<- df_train[ train.index1,]
hold_out_data  <- df_train[-train.index1,]

model <- lm(TARGET_WINS ~ ., data = train_data)

hold_out_data$Prediction <- predict(model,hold_out_data)

cor(hold_out_data$Prediction,hold_out_data$TARGET_WINS)

summary(model)

ggplotly(ggplot(hold_out_data, aes(x=Prediction, y=TARGET_WINS, colour = TARGET_WINS)) +
    geom_point(shape=1) + 
 scale_color_gradient2(low = "blue",
                            high = "red", space = "Lab" )     # Use hollow circles
)

toselect.x <- summary(model)$coeff[-1,4] < 0.05 #

relevant.x <- names(toselect.x)[toselect.x == TRUE]

sig.formula <- as.formula(paste("TARGET_WINS ~",paste(relevant.x, collapse= "+")))

model <- lm(formula = sig.formula, data = train_data)

summary(model)

```

##Look at Variance Inflation Factors