### 2019-09-27 hw2/HW2.Rmd
Initial commit

---
title: "DATA 621 Homework #2"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(caret)
library(pROC)
```

```{r}
source("hw2.R")
df <- read.csv("./data/classification-output-data.csv")
actual <- "class"
predicted <- "scored.class"

cm <- confusion_matrix(df, actual, predicted)
```