### 2019-10-06 hw2/HW2.Rmd
More work on HW2

---
title: "DATA 621 Homework #2"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(caret)
library(pROC)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv("./data/classification-output-data.csv")
actual <- "class"
predicted <- "scored.class"
probability <- "scored.probability"
# Load our functions
source("hw2.R")
```

## Caret's Output

```{r}
confusionMatrix(factor(df[[predicted]]), factor(df[[actual]]))
```

## Our Function's Output

**Confusion Matrix**

```{r}
confusion_matrix(df, actual, predicted) %>%
  kable() %>%
  kable_styling()
```

**Accuracy**: `r accuracy(df, actual, predicted)`

**Classification Error Rate**: `r classification_error_rate(df, actual, predicted)`

**Precision**: `r precision(df, actual, predicted)`

**Sensitivity (Recall)**: `r sensitivity(df, actual, predicted)`

**Specificity**: `r specificity(df, actual, predicted)`

**F1 Score**: `r f1_score(df, actual, predicted)`