### 2019-10-30 hw3/HW3.Rmd
Folded in Corey's work

---
title: "DATA 621 Homework #3"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
library(Amelia)
library(pROC)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv('data/crime-training-data_modified.csv')
evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
```

## Data Exploration

### Are There Missing Values?
First we look at the data to see if any variables have missing data:

```{r}
missmap(df, main = "Missing vs Observed Values")
```

It looks like we have a complete data set.  No need to impute values.

### Splitting the Data
Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
```{r}
set.seed(42)
train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Exploratory Data Analysis
Now, let's look at our training data. By looking at a correlation matrix, we can see which variables may be too correlated to be included together in a model as predictor variables. This will help us later during the model selection process.

```{r}
train %>% 
  cor(.) %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```

Next we will look at each potential predictor and how it is distributed across the target variable.

The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e. blue) and a negative target variable (areas with crime rates below the median, i.e. red). What we are looking for is variables that show way to split data into two groups.

```{r}
for (var in names(train)){
  if(var != "target"){
    plot_df <- train
    plot_df$x <- plot_df[,var]
    p <- ggplot(plot_df, aes(x, color = factor(target))) +
      geom_density() +
      theme_light() +
      ggtitle(var) +
      scale_color_brewer(palette = "Set1") +
      theme(legend.position = "none",
            axis.title.y = element_blank(),
            axis.title.x = element_blank())
    print(p)
  }
}
```

Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.

## Data Preparation

### Feature Engineering
```{r}
library(fuzzyjoin)
density_diff <- function(data,eval_data,var,target,new_column){
  data[[paste0(var,"temp")]] <- data[[var]]
  eval_data[[paste0(var,"temp")]] <- eval_data[[var]]

  #standardize variable between 0 and 1
  data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))
  eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))

  ## Calculate density estimates
  g1 <- ggplot(data, aes(x=data[[var]], group=target, colour=target)) +
    geom_density(data = data) + xlim(min(data[[var]]), max(data[[var]]))
  gg1 <- ggplot_build(g1)
  # construct the dataset
  x <- gg1$data[[1]]$x[gg1$data[[1]]$group == 1]
  y1 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 1]
  y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
  df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), side=(y1<y2), ydiff = y2-y1)
  ##creating the second graph object
  g3 <- ggplot(df2) +
    geom_line(aes(x = x, y = ydiff, colour = side)) +
    geom_area(aes(x = x, y = ydiff, fill = side, alpha = 0.4)) +
    guides(alpha = FALSE, fill = FALSE)

  data$join <- data[[var]]
  df2$join <- df2$x
  temp <- difference_left_join(data,df2, by="join", max_dist =.001)
  means <- aggregate(ydiff ~ temp$join.x, temp, mean)

  colnames(means) <- c("std_var","density_diff")

  new_data <- merge(means,data, by.x ="std_var", by.y =var)
  new_data$std_var <- NULL
  new_data$join <- NULL
  new_data[new_column] <- new_data$density_diff
  new_data$density_diff <- NULL
  ###################same thing with eval data ############################
  eval_data$join <- eval_data[[var]]
  df2$join <- df2$x
  temp <- difference_left_join(eval_data,df2, by="join", max_dist =.001)
  means <- aggregate(ydiff ~ temp$join.x, temp, mean)

  ##new stuff
  colnames(means) <- c("std_var","density_diff")
  #data["key"] <- (var - min(var))/(max(var) - min(var))

  eval_data2 <- merge(means,eval_data, by.x ="std_var", by.y =var)
  eval_data2$std_var <- NULL
  eval_data2$join <- NULL
  eval_data2[new_column] <- eval_data2$density_diff
  eval_data2$density_diff <- NULL

  eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
  eval_data2[[paste0(var,"temp")]] <- NULL

  new_data[[var]] <- new_data[[paste0(var,"temp")]]
  new_data[[paste0(var,"temp")]] <- NULL

  mylist <- list(g1,g3,new_data,eval_data2)
  names(mylist)<- c("dist","dist_diff","new_data_training","new_data_eval")
  return(mylist)
}
```

### Creating Some New Variables

```{r}
str(train)
str(test)
## try it out 
train <- density_diff(train,test,"nox",train$target,"nox_density")$new_data_training
train <- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
train <- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
train <- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
train <- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
train <- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
train <- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
train <- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
train <- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training

test<- density_diff(train,test,"nox",train$target,"nox_density")$new_data_eval
test<- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training

str(train)
str(test)
```

## Model Building

We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.

```{r}
baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
summary(baseline)
test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
                  deviance=baseline$deviance, 
                  r2 = 1 - baseline$deviance/baseline$null.deviance,
                  aic=baseline$aic)
cm
```

Our baseline model is ok with an F1 score of `r cm$byClass[7]`.

Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:

```{r}
fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
summary(fullmodel)
test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
results <- rbind(results,tibble(model = "fullmodel",
                                predictors = 12,F1 = cm$byClass[7],
                                deviance=fullmodel$deviance, 
                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
                                aic=fullmodel$aic))
cm
```

The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.

After some backward elimination of non-significant predictor variables, we arrive at the following model:

```{r}
model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
              family = binomial(link = "logit"), 
              train)
summary(model1)
test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
results <- rbind(results,tibble(model = "model1",
                                predictors = 6,F1 = cm$byClass[7],
                                deviance=model1$deviance, 
                                r2 = 1-model1$deviance/model1$null.deviance,
                                aic=model1$aic))
cm
```

```{r}
density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
summary(density_models)
test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
test$density_models_yhat <- predict(density_models, test)
cm <- confusionMatrix(factor(test$density_models), factor(test$target))
results <- rbind(results,tibble(model = "density models",
                                predictors = 9,
                                F1 = cm$byClass[7],
                                deviance = density_models$deviance,
                                r2 = 1 - density_models$deviance / density_models$null.deviance,
                                aic = density_models$aic))
cm
```

Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.

```{r}
kable(results)
```

Looking these three models together, `model1` looks like the best one. Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.

## Model Selection

### pROC Output

#### Baseline
```{r}
par(pty = "s")
roc(test[["target"]], test[["baseline"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Full Model
```{r}
par(pty = "s")
roc(test[["target"]], test[["fullmodel"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Model1
```{r}
par(pty = "s")
roc(test[["target"]], test[["model1"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Density Model
```{r}
par(pty = "s")
roc(test[["target"]], test[["density_models_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```