### 2019-11-05 final project/Final_Project.Rmd
Update final

---
title: "DATA 621 - Final Project"
author:
  - name: Corey Arnouts
    affiliation: CUNY School of Professional Studies
  - name: Adam Douglas
    affiliation: CUNY School of Professional Studies
  - name: Jason Givens-Doyle
    affiliation: CUNY School of Professional Studies
  - name: Michael Silva
    affiliation: CUNY School of Professional Studies
    email: Michael.Silva39@spsmail.cuny.edu
date: "`r Sys.Date()`"
output:
  html_document: default
  rticles::asa_article:
    citation_package: natbib
    keep_tex: yes
    number_sections: no
    toc: no
keywords: keywordA, keywordB
abstract: Use 250 words or less to summarize your problem, methodology, and major outcomes. \newline This is the second line of abstract.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Describe the background and motivation of your problem.

## Literature Review

Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are. Explain how your investigation is similar or different to the state-of-theart. Please cite the relevant papers where appropriate.

## Methodology

Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.

## Experimentation and Results

Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).

## Discussion and Conclusions

Conclude your findings, limitations, and suggest areas for future work.

## References

Be sure to cite all references used in the report (APA format).

## Appendices

- Supplemental tables and/or figures.
- R statistical programming code.