### 2019-11-08 hw4/HW4_Mike.Rmd
Initial commit of my scratchpad

---
title: "Mike's Scratchpad for DATA 621 Homework #4"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(RColorBrewer)
library(FFTrees)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

## Data Cleanup

The dollar values have dollar signs.  These will be stripped out and converted from a character to numeric.  LOG10 and LN transformations of all dollar values will be preformed.  Categorical data will be translated into dummy variables.

```{r}
df <- read.csv("./data/insurance_training_data.csv")
evaluation <- read.csv("./data/insurance-evaluation-data.csv")

strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}
```

```{r}
to_dummies <- function(df, var, drop = FALSE){
  temp <- data.frame(INDEX = df$INDEX, CAT = df[[var]]) %>%
    mutate(CAT = as.character(CAT),
           VAL = 1) %>%
    spread(CAT, VAL, fill = 0)
  new_names <- c()
  for (n in names(temp)){
    if(n != "INDEX"){
      n <- paste0(var,"_",n)
      n <- gsub(" ", "_", n)
    }
    new_names <- c(new_names, n)
  }
  names(temp) <- new_names

  if(drop){
    drops <- c(var)
    df <- df[ , !(names(df) %in% drops)]
  }

  df %>%
    merge(temp)
}

get_driver_type <- function(x, y){
  if(is.na(x) | is.na(y)){
    return(NA)
  }
  if(y < 40){
    driver_type <- "Young"
  } else if (y < 70){
    driver_type <- "Middle Age "
  } else {
    driver_type <- "Older"
  }
  if( x == "M"){
    return(paste(driver_type, "Male"))
  } else {
    return(paste(driver_type, "Female"))
  }
}

temp <- df %>%
  rowwise() %>%
  mutate(INCOME = strip_dollars(INCOME),
         HOME_VAL = strip_dollars(HOME_VAL),
         BLUEBOOK = strip_dollars(BLUEBOOK),
         OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
  mutate(JOB = ifelse(str_length(JOB) > 0, JOB, NA),
         CAR_AGE = ifelse(str_length(CAR_AGE) > 0, CAR_AGE, NA),
         EDUCATION = ifelse(EDUCATION == "<High_School", "LT_High_School", EDUCATION),
         URBANICITY = ifelse(URBANICITY == "Highly Urban/ Urban", "Urban", "Rural")) %>%
  mutate(DRIVER_TYPE = get_driver_type(SEX, AGE)) %>%
  ungroup() %>%
  na.omit() %>%
  to_dummies("PARENT1", T) %>%
  to_dummies("KIDSDRIV", T) %>%
  to_dummies("HOMEKIDS", T) %>%
  to_dummies("MSTATUS", T) %>%
  to_dummies("SEX", T) %>%
  to_dummies("EDUCATION", T) %>%
  to_dummies("JOB", T) %>%
  to_dummies("CAR_USE", T) %>%
  to_dummies("CAR_TYPE", T) %>%
  to_dummies("RED_CAR", T) %>%
  to_dummies("REVOKED", T) %>%
  to_dummies("MVR_PTS", T) %>%
  to_dummies("URBANICITY", T) %>%
  to_dummies("DRIVER_TYPE", T) %>%
  rowwise() %>%
  mutate(BLUEBOOK = ifelse(BLUEBOOK > 0, BLUEBOOK, 0)) %>%
  # mutate(LOG10_INCOME = log10(INCOME + 1),
  #        LN_INCOME = log(INCOME + 1),
  #        LOG10_HOME_VAL = log10(HOME_VAL + 1),
  #        LN_HOME_VAL = log(HOME_VAL + 1),
  #        LOG10_BLUEBOOK = log10(BLUEBOOK + 1),
  #        LN_BLUEBOOK = log(BLUEBOOK + 1),
  #        LOG10_OLDCLAIM = log10(OLDCLAIM + 1),
  #        LN_OLDCLAIM = log(OLDCLAIM + 1)) %>%
  mutate(PRIOR_CLAIM = ifelse(CLM_FREQ + OLDCLAIM > 0, 1,0)) %>%
  ungroup() 
```

```{r}
set.seed(42)
minority <- nrow(temp[temp$TARGET_FLAG == 1,])
majority <- nrow(temp[temp$TARGET_FLAG == 0,])
diff <- majority - minority
minority_index <- temp[temp$TARGET_FLAG == 1,]$INDEX
over_sample_df <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
  merge(temp, .) %>%
  bind_rows(temp)

train_index <- createDataPartition(over_sample_df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
train <- over_sample_df[train_index,]
test <- over_sample_df[-train_index,]
```

```{r}
fft_train <- train %>%
  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
fft_test <- test %>%
  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
fit <- FFTrees(TARGET_FLAG ~ ., fft_train, fft_test)
plot(fit)
```
The FFTree suggests that past history of filing claims is a good predictor if someone will file a claim in the future.  This could be our baseline model criteria.

Other variables that seems to come up is if the driver is in an urban location and possibly their home value and if it's a commercial vehicle