### 2019-11-27 hw5/HW5.rmd

---
title: "Data-621-Project-5"
author: "Critical Thinking Group 3"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r}
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(purrr)
library(RColorBrewer)
library(ROCR)
library(corrplot)
```

```{r}

df <- read.csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw5/data/wine-training-data.csv")

hold_out_data <- read.csv("https://raw.githubusercontent.com/mikeasilva/data-621-group-projects/master/hw5/data/wine-evaluation-data.csv")
```

```{r}
str(df)
```
#Look at the Distribution of the Target Variable

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = df, aes(x = TARGET)) +
  geom_bar(width = 1, color = 'black',fill = I('orange'))+ggtitle("Number of Wine Cases Purchased")

```

## Training Test Split
```{r}
set.seed(42)
train_index <- createDataPartition(df$TARGET, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

## Lets Look at some decision Trees wooo!

```{r}
source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")

test <- RF_with_Nulls(train,test,"TARGET",.5,10,5,.005,4,1)

```

```{r}

df_complete <- df[complete.cases(df), ]

library(corrplot)
M <- cor(df_complete)
corrplot(M, method = "circle") #plot matrix
library("PerformanceAnalytics")
```

## Analyze Relationship Between Variables

### Strong Relationship between Expert Rating (STARS) and the number of cases purchased
```{r}
# Set color by cond
ggplot(df_complete, aes(x=as.factor(STARS), y=TARGET,color =as.factor(STARS))) + geom_boxplot()+
  labs(title = "Cases of Wine Purchased Compared to Expert Rating", x = "Expert Rating", y = "Cases Purchased", color = "Expert Rating")

```

```{r}
# Set color by cond
ggplot(df_complete, aes(x=as.factor(LabelAppeal), y=TARGET,color =as.factor(LabelAppeal))) + geom_boxplot()+
  labs(title = "Cases of Wine Purchased Compared to Label Appeal", x = "Label Appeal", y = "Cases Purchased", color = "LabelAppeal")

```

## Fitting a Poisson Model to the Data

```{r}
model <- glm(TARGET ~ LabelAppeal + STARS +AcidIndex+VolatileAcidity,
             family = poisson, train)

summary(model)

test$Prediction <- predict.glm(model, newdata=test, type="response")

cor(test$Prediction,test$TARGET)
```